---
title: "Stage 2 - Annotating the data"
author: "Kathryn Steel, Yann Abraham"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
version: 0.1.0
---

```{r setup}
library(targets)

library(tibble)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)

knitr::opts_chunk$set(warning=FALSE,
                      echo=TRUE,
                      fig.width=8,
                      fig.height=6,
                      fig.retina=1,
                      fig.keep='high',
                      fig.align='center',
                      knitr.table.format = "html")

theme_set(theme_light(base_size = 16))
```

```{targets pipeline, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("flowCore","flowWorkspace",
                            "dplyr","tidyr","stringr","forcats","tibble",
                            "ggplot2","Rtsne","Radviz",
                            "kohonen","igraph","rmarkdown"))
## functions
source("scripts/peptidescore_src.R")
```

# Goal

In this step and based on the information collected during Stage1, we annotate samples and markers and extract the intensity matrix to support the analysis.

# Sample annotation

Based on the information collected in Stage1 we prepare the sample annotation target:

```{targets annots}
tar_target(annots, {
  .gs <- load_gs(gs.file)
  .annots <- pData(.gs) %>%
    select(`FCS Filename` = name) %>%
    rownames_to_column("name") %>%
    mutate(Treatment = str_extract(`FCS Filename`,
                                   "CEFT|CEF|TT|HEX|[RC]VIM|[RC]ENO|[RC]AGG|[RC]FIB|[RC]CILP"),
           Individual = str_extract(`FCS Filename`,
                                    "MET[0-9]{3}"),
           Treatment = factor(Treatment),
           Individual = factor(Individual),
           Treatment = fct_relevel(Treatment,"CEF","CEFT","TT","HEX"),
           Sample = paste0(Individual,"_",Treatment),
           Total = str_extract(name,"_[0-9]{4,}"),
           Total = str_replace(Total,"_",""),
           Total = as.numeric(Total))
  .annots
})
```

# Marker annotation

Based on the information collected in Stage1 we prepare the sample annotation target:

```{targets markers}
tar_target(markers,{
  .gs <- load_gs(gs.file)
  as_tibble(pData(parameters(flowData(.gs)[[1]])),
            rownames = "id") %>% 
    mutate(id = paste0(id,"S"),
           type = case_when(
             desc %in% params_ref_markers ~ "reference",
             is.na(desc) ~ "instrument",
             TRUE ~ "functional"
           ),
           desc_orig = desc,
           desc = ifelse(is.na(desc),name,desc),
           isTransformed  = name %in% names(getTransformations(.gs[[1]])))
})
```

We record which marker type will be used for the remainder of the analysis:

```{targets markertype}
tar_target(markertype,"functional")
```


# Intensity matrix

We extract the intensity matrix from the data for the gate of interest and turn the object into a tibble including annotation:

```{targets div.df}
tar_target(div.df,{
  .gs <- load_gs(gs.file)
  
  .mats <- getData(.gs,params_cur_pop)
  .mats <- fsApply(.mats,exprs,simplify = FALSE)
  .mats <- lapply(.mats,function(mat) {
    as_tibble(mat)%>% 
      select(with(markers,name[type!="instrument"])) %>% 
      rename_with(~ with(markers,desc[type!="instrument"]))
  })
  bind_rows(.mats,
            .id = "name") %>% 
    left_join(annots,
              by = "name") %>% 
    group_by(name) %>% 
    mutate(id = paste0(name,seq(length(name))))
})
```

For distance computation we extract the intensity values and turn them into a matrix with rownames corresponding to single cell identifier:

```{targets div-mat}
tar_target(div.mat,{
  .mat <- div.df %>% 
    ungroup() %>% 
    select(with(markers,
                unname(desc[type==markertype]))) %>% 
    as.matrix()
  rownames(.mat) <- div.df$id
  .mat
})
```

# Run the pipeline

```{r}
tar_make()
```
