---
title: "Stage 6 - Exploratory Analysis"
author: "Kathryn Steel, Yann Abraham"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
version: 0.1.0
---

```{r setup}
library(targets)
library(tarchetypes)

library(tibble)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)

knitr::opts_chunk$set(warning=FALSE,
                      echo=TRUE,
                      fig.width=8,
                      fig.height=6,
                      fig.retina=1,
                      fig.keep='high',
                      fig.align='center',
                      knitr.table.format = "html")

theme_set(theme_light(base_size = 16))
```

```{targets pipeline, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("flowCore","flowWorkspace",
                            "dplyr","tidyr","stringr","forcats","tibble",
                            "ggplot2","Rtsne","Radviz",
                            "kohonen","igraph","rmarkdown"))

## functions
source("scripts/peptidescore_src.R")
```

# Goal

Compute a few useful statistics and switch to interactive exploratory analysis.

# Community size

To compute the community size per sample, we start by specifying the variables of interest that should be used for the grouping:

```{targets params_vars}
tar_target(params_vars,
           c("Treatment","Individual"))
```

Next we compute the community size and transform it into a *clr* value:

```{targets div.clr, tar_interactive=FALSE}
tar_target(div.clr,{
  N <- length(div.groups)
  
  div.df %>% 
    ungroup() %>% 
    mutate(Community = factor(membership(div.groups)[id])) %>% 
    count(across(all_of(c(params_vars,"Community")))) %>%
    complete(Community,
             nesting(select(.,all_of(params_vars))),
             fill=list(n=0)) %>% 
    group_by(across(all_of(params_vars))) %>%
    mutate(pc = n/sum(n),
           minim = 1/sum(n,1),
           m = N-sum(n!=0),
           ts = m*minim,
           pc = if_else(n==0,
                        minim,
                        pc - pc*ts),
           clr = log(pc/exp(mean(log(pc))))) %>% 
    select(-minim,-m,-ts)
})
```

# Biplot

To compute a biplot representation of the community contributions per sample, we first set the alpha parameters, which describes the relative contribution of features and samples to the projection; use 1 to focus on features, 0 to focus on samples:

```{targets params_alpha}
tar_target(params_alpha,
           0.5)
```

We can then compute the `svd` representation of the *clr* matrix:

```{targets div.svd, tar_interactive=FALSE}
tar_target(div.svd,{
  .vars <- div.clr %>% 
    ungroup() %>% 
    unite("Sample",
          all_of(params_vars),
          sep = "_",
          remove = FALSE)
  
  .log.pc <- .vars %>% 
    select(Sample,Community,pc) %>% 
    spread(Community,pc)
  .rows <- .log.pc[,1,drop=TRUE]
  .log.pc <- as.matrix(.log.pc[,-1])
  
  ## double centered svd
  .z <- log(.log.pc)
  .z.avg <- mean(.z)
  .z.row <- rowMeans(.z)
  .z.col <- colMeans(.z)
  .z <- sweep(.z,1,.z.row,`-`)
  .z <- sweep(.z,2,.z.col,`-`)
  .z <- .z+.z.avg
  .svd <- svd(.z)
  
  ## biplot
  .f <- .svd$u %*% diag(.svd$d^params_alpha)
  .f <- .f[,seq(2)]
  .f <- tibble(x = .f[,1],
               y = .f[,2],
               Sample = .rows) %>% 
    left_join(.vars %>% 
                select(all_of(c("Sample",params_vars))),
              by = "Sample")
  
  .g <- .svd$v %*% diag(.svd$d^(1-params_alpha))
  .g <- .g[,seq(2)]
  if(params_alpha==0) {
    .g <- .g/((length(.rows)-1)^0.5)
  }
  .g <- tibble(x = .g[,1],
               y = .g[,2],
               Community = colnames(.log.pc))
  list(bip_F = .f,
       bip_G = .g)
})
```

# Reports

Here we can specify which additional reports should be run using the output of the pipeline:

```{targets explore}
tarchetypes::tar_render(explore,
                        "reports/_explore.Rmd",
                        params = list(cache = tar_config_get("store")))
```

# Run the pipeline

```{r}
tar_make()
```
