---
title: "Stage 4 - Find optimal k"
author: "Kathryn Steel, Yann Abraham"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
version: 0.1.0
---

```{r setup}
library(targets)

library(tibble)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)

knitr::opts_chunk$set(warning=FALSE,
                      echo=TRUE,
                      fig.width=8,
                      fig.height=6,
                      fig.retina=1,
                      fig.keep='high',
                      fig.align='center',
                      knitr.table.format = "html")

theme_set(theme_light(base_size = 16))
```

```{targets pipeline, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("flowCore","flowWorkspace",
                            "dplyr","tidyr","stringr","forcats","tibble",
                            "ggplot2","Rtsne","Radviz",
                            "kohonen","igraph","rmarkdown"))

## functions
source("scripts/peptidescore_src.R")
```

# Goal

Based on the distances we computed we estimate the optimal *k* value to represent the data as a graph.

# Find optimal *k*

To identify the optimal *k* we suggest starting from the square root of the number of cells, then exploring a number of critical value around that one.

We start by specifying a range of *k* values to test:

```{targets ks}
tar_target(ks,
           seq(100,200,by = 25))
```

Next we compute the corresponding graphs, and extract the communities using the Louvain algorithm

```{targets comms, tar_interactive=FALSE}
tar_target(comms,{
  .adj <- div.dist %>% 
    filter(rweight<=ks)
  .graph <- graph_from_data_frame(.adj,
                                  directed=FALSE)
  cluster_louvain(.graph)
},
pattern = map(ks)
)
```

Finally we extract the following summary statistics:

  * Number of communities
  * Modularity
  * Gini coefficient of community size

```{targets mods, tar_interactive=FALSE}
tar_target(mods,{
  ## compute a gini coefficient for community size
  .sizes <- sizes(comms)
  .n <- length(.sizes)
  
  tibble(k = ks,
         modularity = modularity(comms),
         Gini = sum(outer(.sizes,
                          .sizes,
                          FUN=function(x,y){
                            abs(x-y)})) / (2 * .n * sum(.sizes)),
         N = length(comms))
},
pattern = map(ks,comms))
```

We visualize the different parameters in the following plot:

```{targets show_modularity, tar_interactive=FALSE}
tar_target(show.mods,{
  mods %>% 
    mutate(k = factor(as.numeric(k))) %>% 
    ggplot(aes(x = N,
               y = modularity))+
    geom_path(size = 1.5,
              col = "grey80")+
    geom_text(aes(label = k,
                  size = Gini))
})
```

# Run the pipeline

```{r}
tar_make()
```

# Show the results

```{r}
tar_read(show.mods)
```

We select *k* as the value where:

  * Number of communities is maximum
  * Modularity is maximum
  * Gini coefficient of community size is maximum
