---
title: "Stage 3 - Computing the distance"
author: "Kathryn Steel, Yann Abraham"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
version: 0.1.0
---

```{r setup}
library(targets)

library(tibble)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)

knitr::opts_chunk$set(warning=FALSE,
                      echo=TRUE,
                      fig.width=8,
                      fig.height=6,
                      fig.retina=1,
                      fig.keep='high',
                      fig.align='center',
                      knitr.table.format = "html")

theme_set(theme_light(base_size = 16))
```

```{targets pipeline, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("flowCore","flowWorkspace",
                            "dplyr","tidyr","stringr","forcats","tibble",
                            "ggplot2","Rtsne","Radviz",
                            "kohonen","igraph","rmarkdown"))

## functions
source("scripts/peptidescore_src.R")
```

# Goal

Now that the data is loaded we start by computing distance between cells using a greedy approach: rather than computing all distances, including ones that would be irrelevant to the construction of the graph later on, we use a hash table to compute distances only between close neighbors

# Parameters

The distance computation uses all markers of type **functional** in the dataset; to efficiently compute a hash table we use **S**elf **O**rganizing **M**aps to quickly identify cells that are close to one another. We start by defining a few key parameters:

    - N: the total number of points
    - B: the number of points to sample and the maximum number of points per hash
    - k: the size of the SOM grid
    - L: the number of hash tables
    - K: the maximum number of nearest neighbors

```{targets simhash_params}
tar_target(simhash_params,
           list(N = nrow(div.df),
                B = 2000,
                k = 2,
                L = 6,
                K = 250))
```

# Distance estimation

First we compute the SOM grid:

```{targets som.grid}
tar_target(som.grid,
           somgrid(xdim = simhash_params$k, 
                       ydim = simhash_params$k, 
                       topo = 'hexagonal', 
                       toroidal = T))
```

Next we compute the hash tables:

```{targets hash.table}
tar_target(hash.table,
           do.hash.table(mat = div.mat,
                         grid = som.grid,
                         params = simhash_params))
```

In the next step we compute **cosine** similarities for all hash tables, and all buckets; the similarities are then aggregated retaining only the largest ones, and the ranks are computed on the result:

```{targets div.dist}
tar_target(div.dist,
           do.hash.dist(mat = div.mat,
                        htable = hash.table,
                        params = simhash_params))
```

# Run the pipeline

```{r}
tar_make()
```
