---
title: "Stage 5 - Build the graph"
author: "Kathryn Steel, Yann Abraham"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
version: 0.1.0
---

```{r setup}
library(targets)

library(tibble)
library(dplyr)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)

knitr::opts_chunk$set(warning=FALSE,
                      echo=TRUE,
                      fig.width=8,
                      fig.height=6,
                      fig.retina=1,
                      fig.keep='high',
                      fig.align='center',
                      knitr.table.format = "html")

theme_set(theme_light(base_size = 16))
```

```{targets pipeline, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("flowCore","flowWorkspace",
                            "dplyr","tidyr","stringr","forcats","tibble",
                            "ggplot2","Rtsne","Radviz",
                            "kohonen","igraph","rmarkdown"))

## functions
source("scripts/peptidescore_src.R")
```

# Goal

Based on our estimation of the optimal *k* value we model the data as a graph and compute a few projections.

# Specify *k* and create the kNN graph

We define the optimal k as a parameter in the following block:

```{targets params_k}
tar_target(params_k,150)
```

Then build the graph accordingly:

```{targets div.graph, tar_interactive=FALSE}
tar_target(div.graph,{
  ## build the adjacency matrix
  .adj <- div.dist %>% 
    filter(rweight<=params_k)
  
  ## build the graph
  graph_from_data_frame(.adj,
                        directed=FALSE)
})
```

And finally extract the communities:

```{targets div.groups, tar_interactive=FALSE}
tar_target(div.groups,
           cluster_louvain(div.graph))
```

# Compute projections

## *t*-SNE

Rather than creating a projection based on the features, we will create a projection based on the geodesic distance between cells:

```{targets div.graph.dist, tar_interactive=FALSE}
tar_target(div.graph.dist,
           distances(div.graph,
                     weights = NA))
```

We can then use *t*-SNE to project the graph:

```{targets div.graph.tsne, tar_interactive=FALSE}
tar_target(div.graph.tsne,
           Rtsne(as.dist(div.graph.dist), 
                 perplexity=50, 
                 check_duplicates = FALSE,
                 is_distance = TRUE))
```

And turn the results into a `ggplot2` object:

```{targets div.graph.tsne.proj}
tar_target(div.graph.tsne.proj,{
  as_tibble(div.graph.tsne$Y,
          .name_repair = "universal",
          quiet = TRUE) %>% 
  rename(tx = `...1`,
         ty = `...2`) %>% 
  mutate(name = rownames(div.graph.dist)) %>% 
  left_join(div.df %>% 
              ungroup() %>% 
              mutate(Community = factor(membership(div.groups)[id])),
            by = c("name" = "id")) %>% 
  ggplot(aes(x = tx,
             y =ty))
})
```

## Graphviz

Use Graphviz to project the graph in the context of the features used to build it:

```{targets gv, tar_interactive=FALSE}
tar_target(gv,{
  .scaled <- apply(div.mat,2,do.L)
  .gv <- do.optimGraphviz(.scaled,
                          graph = div.graph)
  rownames(.gv) <- colnames(.scaled)
  .gv
})
```

Once channel order is optimized we can create the projection:

```{targets div.graph.gv, tar_interactive=FALSE}
tar_target(div.graph.gv,{
  do.radviz(div.df %>% 
              ungroup() %>% 
              mutate(Community = factor(membership(div.groups)[id])),
            springs = gv,
            trans = do.L)
})
```

# Run the pipeline

```{r}
tar_make()
```
