---
title: "Stage 1 - Preparing the data"
author: "Kathryn Steel, Yann Abraham"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
version: 0.1.0
---

```{r setup}
library(targets)

library(flowCore)
library(ncdfFlow)
library(flowWorkspace)

library(openCyto)
library(CytoML)


library(tidyverse)

knitr::opts_chunk$set(warning=FALSE,
                      echo=TRUE,
                      fig.width=8,
                      fig.height=6,
                      fig.retina=1,
                      fig.keep='high',
                      fig.align='center',
                      knitr.table.format = "html")

theme_set(theme_light(base_size = 16))
```

```{targets pipeline, tar_globals = TRUE}
options(tidyverse.quiet = TRUE)
tar_option_set(packages = c("tidyverse"))
## functions
source("scripts/peptidescore_src.R")
```

# Goal

Load & parse the data, save it & set the first few parameters.

# Parse the FlowJo workspace into a `GatingSet`

We start by setting a few local parameters:

 - *ws.file* where to find the FlowJo file
 - *output.dir* where to save the `GatingSet`
 - *gs.id* the name of the `GatingSet`

```{r local-parameters}
input.dir <- "input"
ws.file <- "input/workspace.wsp"
output.dir <- "output"
gs.id <- "test_gs"
```

Next we load the data into a `GatingSet`:

```{r load-data}
unlink(file.path(output.dir,gs.id),
       recursive = TRUE)
# ws <- open_flowjo_xml(ws.file)
# gs <- parseWorkspace(ws,name = "gating")
gs_files <- list.files(path = input.dir,
                       pattern = "fcs",
                       full.names = TRUE)
fs <- read.ncdfFlowSet(files = gs_files)
gs <- GatingSet(fs)
```

After loading we can list the existing populations, list the existing markers and review the available sample annotation.

# Sample annotation

```{r show-samples}
pData(gs)
```

# Marker annotation

```{r show-markers}
pData(parameters(gs_cyto_data(gs)[[1]]))
```

Based on this information we define the markers that will be excluded from the analysis:

```{targets params_ref_markers}
tar_target(params_ref_markers,
           c("LD","CD8","CFSE","GPR56","CD39"))

```

# Populations

The dataset has already been gated so we start from root:

```{targets params_cur_pop}
tar_target(params_cur_pop,"root")
```

# Save the `GatingSet`

We can now save the `GatingSet` and save the corresponding folder as a target:

```{r save-gs}
save_gs(gs,
        file.path(output.dir,gs.id))
cat("Copy the following value to the next `tar_target` block:\n",
    "\t",file.path(output.dir,gs.id),"\n")
```

```{targets gs.file}
tar_target(gs.file,
           "output/test_gs")
```

# Run the pipeline

```{r}
tar_make()
```
