---
title: "PeptideScore Results"
output:
  html_notebook:
    df_print: paged
    code_folding: show
    toc: yes
    toc_float: yes
    toc_depth: 4
---

# loading the packages

```{r setup}
library(targets)

library(knitr)

library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(igraph)
library(Radviz)

theme_set(theme_light(base_size = 16))
tar_config_set(store = "/projects/cytofResults/collaborations/RTCure/analysis/ayann/WP4/PeptideScore/_targets/")
```

# loading the data

```{r data}
tar_load(markers)
tar_load(div.df)
tar_load(div.groups)
tar_load(div.clr)
tar_load(div.svd)
tar_load(params_alpha)
tar_load(div.graph.gv)
tar_load(div.graph.tsne)
```

# graph structure {.tabset}

## graphviz

```{r}
community.cols <- setNames(
  scales::hue_pal()(length(div.groups)),
  seq(length(div.groups))
)
plot(Radviz::rescalePlot(div.graph.gv,0.6))+
  geom_point(aes(color = Community))+
  scale_color_manual(values = community.cols)
```

## *t*-SNE

```{r}
bind_cols(div.df %>% 
            ungroup() %>% 
            mutate(Community = factor(membership(div.groups)[id])),
          as_tibble(div.graph.tsne$Y,
                    .name_repair = "universal",
                    quiet = TRUE) %>% 
            rename(tx = `...1`,
                   ty = `...2`)) %>% 
  ggplot(aes(x = tx,
             y =ty))+
  geom_point(aes(color = Community))+
  scale_color_manual(values = community.cols)
```

# community to treatment {.tabset}

## linechart

```{r}
div.df %>% 
  ungroup() %>% 
  mutate(Community = factor(membership(div.groups)[id])) %>% 
  count(Community,Treatment) %>% 
  group_by(Treatment) %>% 
  mutate(pc = n/sum(n)) %>% 
  ggplot(aes(x = Community,
             y = pc))+
  geom_line(aes(group = Treatment),
            color = "grey80")+
  geom_point(aes(color = Treatment),
             size = 3)+
  scale_y_continuous(labels = scales::percent_format())
```

## stacked barplot

```{r}
div.df %>% 
  ungroup() %>% 
  mutate(Community = factor(membership(div.groups)[id])) %>% 
  count(Community,Treatment) %>% 
  group_by(Treatment) %>% 
  mutate(pc = n/sum(n)) %>% 
  ggplot(aes(x=Treatment, y=pc, fill=Community)) +
  geom_bar(stat="identity")+
  theme(axis.text.x=element_text(angle=45,hjust=1))
```

# patient-level community per treatment {.tabset}

```{r, results="asis"}
patient.cols <- setNames(
  scales::hue_pal()(nlevels(div.df$Individual)),
  levels(div.df$Individual)
)
ksink <- lapply(levels(div.df$Treatment),function(cur.cond) {
  cat("##",cur.cond,"\n\n")
  cur.plot <- div.clr %>% 
    filter(Treatment==cur.cond)%>%
    group_by(Individual) %>% 
    mutate(pc = n/sum(n)) %>% 
    ggplot(aes(x = Community,
               y = pc))+
    geom_line(aes(group = Individual),
              color = "grey80")+
    geom_point(aes(color = Individual),
               size = 3)+
    scale_color_manual(values = patient.cols)+
    scale_y_continuous(labels = scales::percent_format())
  print(cur.plot)
  cat("\n\n")
})
```

# community composition {.tabset}

First we need to summarize the data by community & patient:

```{r}
div.summ <- div.df %>% 
  ungroup() %>% 
  mutate(Community = factor(membership(div.groups)[id])) %>% 
  group_by(Community,Individual) %>% 
  summarise_at(vars(one_of(with(markers,desc[type=="functional"]))),
               median) %>% 
  gather("Channel","value",
         one_of(with(markers,desc[type=="functional"]))) %>% 
  left_join(div.clr %>%
              filter(Treatment!="CD3") %>%
              group_by(Individual) %>%
              summarize(n = sum(n),
                        .groups = "drop_last") %>%
              distinct(Individual, n) %>%
              mutate(ref = median(n),
                     n_gauss = exp((2*ref^2)/-n^2)),
            by = "Individual")
```

For patient-level data, the points are sized after the number of cells detected in each patient.

## overall

```{r}
div.df %>% 
  ungroup() %>% 
  mutate(Community = factor(membership(div.groups)[id])) %>% 
  group_by(Community) %>% 
  summarise_at(vars(one_of(with(markers,desc[type=="functional"]))),
               median) %>% 
  gather("Channel","value",
         one_of(with(markers,desc[type=="functional"]))) %>% 
  ggplot(aes(x = Channel,
             y = value))+
  geom_line(aes(group = Community),
            color = "grey80")+
  geom_point(aes(color = Community),
             size = 3)+
  scale_color_manual(values = community.cols)+
  theme(axis.text.x=element_text(angle=45,hjust=1))
```

## per patient {.tabset}

```{r, results="asis"}
ksink <- lapply(levels(div.summ$Community),function(cur.comm) {
  cat("###",cur.comm,"\n\n")
  cur.plot <- div.summ %>% 
    filter(Community==cur.comm) %>% 
    ggplot(aes(x = Channel,
               y = value))+
    geom_line(aes(group = Individual),
              color = "grey80")+
    geom_point(aes(color = Individual,
                   size = n_gauss))+
    scale_size_area()+
    guides(size = "none")+
    theme(axis.text.x=element_text(angle=45,hjust=1))
  print(cur.plot)
  cat("\n\n")
})
```

# biplots

Biplots are generated after [Aitchinson and Greenacre (2002)](https://doi.org/10.1111/1467-9876.00275) using `alpha`=`r params_alpha`.

## overall

```{r}
ggplot()+
  geom_segment(data = div.svd[["bip_G"]],
               aes(x = 0,
                   y = 0,
                   xend = x,
                   yend = y))+
  geom_text(data = div.svd[["bip_G"]],
            aes(x = x,
                y = y,
                label = Community),
            color = "blue2")+
  geom_point(data = div.svd[["bip_F"]],
             aes(x = x,
                 y = y,
                 color = Treatment),
             size = 3)+
  scale_x_continuous(expand = c(0,0.5))
```

## treatment {.tabset}

```{r, results="asis"}
ksink <- lapply(levels(div.df$Treatment),function(cur.cond) {
  cat("###",cur.cond,"\n\n")
  cur.plot <- ggplot()+
    geom_segment(data = div.svd[["bip_G"]],
                 aes(x = 0,
                     y = 0,
                     xend = x,
                     yend = y))+
    geom_text(data = div.svd[["bip_G"]],
              aes(x = x,
                  y = y,
                  label = Community),
              color = "blue2")+
    geom_point(data = div.svd[["bip_F"]] %>% 
                 filter(Treatment==cur.cond),
               aes(x = x,
                   y = y,
                   color = Individual),
               size = 3)+
    scale_x_continuous(expand = c(0,0.5))
  print(cur.plot)
  cat("\n\n")
})
```

## individuals {.tabset}

```{r, results="asis"}
ksink <- lapply(levels(div.df$Individual),function(cur.patient) {
  cat("###",cur.patient,"\n\n")
  cur.plot <- ggplot()+
    geom_segment(data = div.svd[["bip_G"]],
                 aes(x = 0,
                     y = 0,
                     xend = x,
                     yend = y))+
    geom_text(data = div.svd[["bip_G"]],
              aes(x = x,
                  y = y,
                  label = Community),
              color = "blue2")+
    geom_point(data = div.svd[["bip_F"]] %>% 
                 filter(Individual==cur.patient),
               aes(x = x,
                   y = y,
                   color = Treatment),
               size = 3)+
    scale_x_continuous(expand = c(0,0.5))
  print(cur.plot)
  cat("\n\n")
})
```

